Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-12 Cost_manuscript.Rmd
adding bibliography with settings

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey, Gary Kane, & Joseph T. McGuire"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
#bibliography: mPFC_paper.bib
---

@@

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes (Kacelnik and Marsh, 2002; Mochon, Norton, and Ariely, 2012; Hernandez Lallement et al., 2014), and positive outcomes imbue value into their preceding effortful behavior (i.e. learned industriousness, Eisenberger, 1992). Though intuitively valid and empirically supported, this explanation leaves open questions about what features of an effortful task can make it attractive in its own right, and how contextual factors mold its perceived costs (Inzlicht, Shenhav, and Olivola, 2018). Understanding these intrinsic qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

  What is often missing from existing accounts is the common case of choosing among types of behavioral demands. Studies have traditionally compared the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior (Ainslie, 1975; Frederick et al., 2002; Kable and Glimcher, 2007; Kool et al., 2010; Westbrook, Kester, and Braver, 2013; Green, Fristoe, and Myerson, 1994). Although this line of work suggests that discounting functions are similar for delay and effort (Prevost et al., 2010; Massar et al., 2015; Seaman et al., 2018), there is ongoing debate about the shape of the discounting function across effort types (Arulpragasam et al., 2018; Prevost et al., 2010; Chong et al., 2016; Klein-Flugge et al., 2016; Kool and Botvinick, 2018), and discount rates do not always correlate across cost domains (Westbrook and Braver, 2013; Seaman et al., 2018). However, sporadic preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context (Kool & Botvinick, 2018). For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting (e.g. taking the cumbersome way home rather than waiting for a delayed train, even if the resulting time of arrival is the same). Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available (Olivola and Shafir, 2013). Recent work has proposed that the subjective costs produced by delays and cognitive effort is modulated by the possibility of missing out on better concurrent alternatives, namely the opportunity cost of time (OC; Fawcett et al., 2012; Kurzban et al, 2013; Otto and Daw, 2019). This evidence sets the stage for examining what people would actually select when given the option for multiple types of demands, including the unexplored question of deciding between cognitive and physical effort (Inzlicht, Shenhav, and Olivola, 2018).

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Stephens & Krebs, 2019). Foraging has recently attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors (Hayden 2018; Mobbs, Trimmer, Blumstein, & Dayan, 2018), and offers a number of desirable qualities. First, it allows for the estimation of optimal, reward maximizing choice patterns based on the opportunity cost of time (Charnov, 1975; Krebs et al., 1977). We can therefore interpret observed deviations in acceptance rates from such optimal behavior in addition to comparing them among demands. Second, the sequential nature of foraging allows individuals to emphasize different elements of the decision context. For example, the cost (or value) of effort or delay could globally depend on the availability of alternative demands (as suggested above), impacting the perceived opportunity cost of time. Alternatively, it could be intrinsic and robust to such manipulation, and/or depend on task-specific factors that are adjusted once individuals experience different forms of behavioral costs. Foraging rests on an abundant modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019).

  In this study, we tested the hypothesis that preferences would vary depending on whether individuals face a single or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). We explored this question in two foraging experiments: a between-subjects design in which participants foraged for rewards while performing a single demand, and a within-subjects extension that had individuals forage for demands in addition to rewards. We leveraged the ecological and computational richness of foraging paradigms to provide new insights on a number of open questions (Inzlicht, Shenhav, and Olivola, 2018). First, the existence of a well-defined optimal choice strategy helped us identify when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands allowed us to asses whether their cost was intrinsic or relative. Finally, confronting participants with both mental and physical effort let us probe whether these preferences were domain general or task-specific, and in the case of the latter, isolate factors that influenced preferences beyond the often-cited associative relationship between effort and outcome.

@@

- The observed choice dynamics ruled out a number of explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. Effort preferences due to boredom would not have shown the global acceptance decline and convergence in the within-subjects experiment, which also discards the possibility that all our participants were high on need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Even though the reciprocal relationship between effort and outcome can often explain preferences for effort (learned industriousness, Inzlicht2018), the observed differences in acceptances between effort types and the eventual convergence of rates in the within-subjects study cannot be explained by reward alone, as demands were rewarded at the same rate. Instead, we proposed that subjective time sensitivity has been shown to play a role in temporal discounting, improving on classic hyperbolic fits (Mckelachar....). Our findings do not discard the influence of these elements in daily effort evaluations. It supplements our understanding of what drives peoples preferences.

@@

- Control/Exp Design: in demand experiments, chosen options are often carried out after the experiment, but there are known discrepancies between expected and obtained value that could affect repeated decisions in real life. Alt from grant: Investigations of the pursuit and receipt stages have been comparatively limited. Many studies have sought to isolate the evaluation stage by presenting subjects with choices whose results do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [5,11,15,16,19]. However, discrepancies between expected and experienced value are well known [20], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards [3,21,22].

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey, Gary Kane, & Joseph T. McGuire"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
bibliography: cost_paper.bib
---

@@

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards [@Westbrook2015; @Shenhav2017]. Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort [@Ainslie1975; @Hull1943; @Kool2018; @Kool2010; @Treadway2009; @Westbrook2013]. However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes [@Kacelnik2002; @Mochon2012; @HernandezLallement2014], and positive outcomes imbue value into their preceding effortful behavior [i.e. learned industriousness; @Eisenberger1992]. Though intuitively valid and empirically supported, this explanation leaves open questions about what features of an effortful task can make it attractive in its own right, and how contextual factors mold its perceived costs [@Inzlicht2018]. Understanding these intrinsic qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

  What is often missing from existing accounts is the common case of choosing among types of behavioral demands. Studies have traditionally compared the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [@Ainslie1975; @Frederick2002; @Kable2007; @Kool2010; @Westbrook2013; @Green1994). Although this line of work suggests that discounting functions are similar for delay and effort [@Prevost2010; @Massar2015; @Seaman2018], there is ongoing debate about the shape of the discounting function across effort types [@Arulpragasam2018; @Prevost2010; @Chong2016; @Klein-Flugge2016; @Kool2018], and discount rates do not always correlate across cost domains [@Westbrook2013; @Seaman2018]. However, sporadic preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context [@Kool2018]. For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting (e.g. taking the cumbersome way home rather than waiting for a delayed train, even if the resulting time of arrival is the same). Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available [@Olivola2013]. Recent work has proposed that the subjective costs produced by delays and cognitive effort is modulated by the possibility of missing out on better concurrent alternatives, namely the opportunity cost of time [OC; @Fawcett2012; @Kurzban2013; @Otto2019]. This evidence sets the stage for examining what people would actually select when given the option for multiple types of demands, including the unexplored question of deciding between cognitive and physical effort [@Inzlicht2018; @Kool2018].

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead [@Stephens2019]. Foraging has recently attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors [@Hayden2018; @Mobbs2018], and offers a number of desirable qualities. First, it allows for the estimation of optimal, reward maximizing choice patterns based on the opportunity cost of time [@Charnov1975; @Krebs1977]. We can therefore interpret observed deviations in acceptance rates from such optimal behavior in addition to comparing them among demands. Second, the sequential nature of foraging allows individuals to emphasize different elements of the decision context. For example, the cost (or value) of effort or delay could globally depend on the availability of alternative demands (as suggested above), impacting the perceived opportunity cost of time. Alternatively, it could be intrinsic and robust to such manipulation, and/or depend on task-specific factors that are adjusted once individuals experience different forms of behavioral costs. Foraging rests on an abundant modeling history that can help to computationally capture such emphases [@Stephens2019].

  In this study, we tested the hypothesis that preferences would vary depending on whether individuals face a single or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). We explored this question in two foraging experiments: a between-subjects design in which participants foraged for rewards while performing a single demand, and a within-subjects extension that had individuals forage for demands in addition to rewards. We leveraged the ecological and computational richness of foraging paradigms to provide new insights on a number of open questions [@Inzlicht2018]. First, the existence of a well-defined optimal choice strategy helped us identify when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands allowed us to asses whether their cost was intrinsic or relative. Finally, confronting participants with both mental and physical effort let us probe whether these preferences were domain general or task-specific, and in the case of the latter, isolate factors that influenced preferences beyond the often-cited associative relationship between effort and outcome.

@@

- The observed choice dynamics ruled out a number of explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. Effort preferences due to boredom would not have shown the global acceptance decline and convergence in the within-subjects experiment, which also discards the possibility that all our participants were high on need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Noise also discarded (Hayden paper), as values on each end of the reward range moved little over time. Even though the reciprocal relationship between effort and outcome can often explain preferences for effort (learned industriousness, Inzlicht2018), the observed differences in acceptances between effort types and the eventual convergence of rates in the within-subjects study cannot be explained by reward alone, as demands were rewarded at the same rate. Instead, we proposed that subjective time sensitivity has been shown to play a role in temporal discounting, improving on classic hyperbolic fits (Mckelachar....). Our findings do not discard the influence of these elements in daily effort evaluations. It supplements our understanding of what drives peoples preferences.

@@

- Control/Exp Design: in demand experiments, chosen options are often carried out after the experiment, but there are known discrepancies between expected and obtained value that could affect repeated decisions in real life. Alt from grant: Investigations of the pursuit and receipt stages have been comparatively limited. Many studies have sought to isolate the evaluation stage by presenting subjects with choices whose results do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [5,11,15,16,19]. However, discrepancies between expected and experienced value are well known [20], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards [3,21,22]. Relying on a foraging paradigm allowed us to control for these and other variables...then control/exp design?
